CAKE: Classification \textit{avec} Kernel Embeddings \\ using Rademacher Complexity Bounds

### 2018-03-29 AISTATS2018/cake_aistats2018.tex
Previous changes

\usepackage{glossaries}
\usepackage{acronym}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[colorlinks=true,linkcolor=green,citecolor=cyan]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

@@

		One notable success story in this domain are \gls{GPs}, which utilizes the marginal likelihood for hyperparameter learning. The marginal likelihood arises from its Bayesian formulation, and exhibits certain desirable properties -- in particular, the ability to automatically balance between data fit and model complexity. On the other hand, \glspl{CME} are not necessarily Bayesian, and hence they do not benefit from a natural marginal likelihood formulation, yet such a balance is critical when generalizing the model beyond unseen examples.

@@

		Let $p_{c}(x) := \mathbb{P}[Y = c | X = x]$ be the \textit{decision probability function} for class $c \in \mathbb{N}_{m}$, which is the probability of the class label $Y$ being $c$ when the example $X$ is $x$. Note that there are no restrictions on the input domain $\mathcal{X}$ as long as a kernel $k$ can be defined on it. For example, $x$ could be a continuous variable in $\mathcal{X} = \mathbb{R}^{d}$. We begin by writing this probability as an expectation of indicator functions,
		\begin{equation}
		p_{c}(x) := \mathbb{P}[Y = c | X = x] = \mathbb{E}[\mathbb{1}_{c}(Y) | X = x].
		\label{eq:decision_probability}
		\end{equation}	
		With $\mathbb{1}_{c} \in \mathcal{H}_{\delta}$, we let $g = \mathbb{1}_{c}$ in \eqref{eq:empirical_conditional_expectation} and $\bvec{1}_{c} := \{\mathbb{1}_{c}(y_{i})\}_{i = 1}^{n}$ to estimate the right hand side of \eqref{eq:decision_probability} by
		\begin{equation}
		\hat{p}_{c}(x) = f_{c}(x) := \bvec{1}_{c}^{T} (K + n \lambda I)^{-1} \bvec{k}(x).
		\label{eq:empirical_decision_probability}
		\end{equation}
		Let $\bvec{Y} := \begin{bmatrix} \bvec{1}_{1} & \bvec{1}_{2} & \cdots & \bvec{1}_{m} \end{bmatrix} \in \{0, 1\}^{n \times m}$ be the one hot encoded labels of $\{y_{i}\}_{i = 1}^{n}$. The vector of empirical decision probabilities over the classes $c \in \mathbb{N}_{m}$ is then
		\begin{equation}
		\hat{\bvec{p}}(x) = \bvec{f}(x) := \bvec{Y}^{T} (K + n \lambda I)^{-1} \bvec{k}(x) \in \mathbb{R}^{m}.
		\label{eq:empirical_decision_probability_vector}
		\end{equation}
		Since $\mathcal{U} = \hat{\mathcal{U}}_{Y | X}$ \eqref{eq:empirical_conditional_embedding} is the solution to a regularized least squares problem in the RKHS from $k(x, \cdot) \in \mathcal{H}_{k}$ to $l(y, \cdot) \in \mathcal{H}_{l}$ \citep{lever2012conditional}, \glspl{CME} can be seen as \glspl{KRR} in the \gls{RKHS}. In this case, because $\mathcal{Y} = \mathbb{N}_{m}$ is discrete, $\mathcal{H}_{\delta}$ can be identified with $\mathbb{R}^{m}$. As a result, for discrete $\mathcal{Y} = \mathbb{N}_{m}$ the rows of the \gls{MCE} can also be seen $m$ \glspl{KRR} \citep{friedman2001elements} on binary $\{0, 1\}$-targets, where they all share the same input kernel. We now verify that the empirical decision probabilities \eqref{eq:empirical_decision_probability} do converge to the population decision probability.

@@

		This also allows \glspl{MCE} to be naturally applied to perform probabilistic classification in multiclass settings with categorical targets. In contrast, \glspl{SVC} do not naturally produce probabilities, while \glspl{GPC} require posterior approximations. Furthermore, as \glspl{SVC} and \glspl{GPC} are original binary classifiers, multiclass extensions often employ the \gls{OVA} or \gls{OVO} scheme \citep{aly2005survey}. Instead, fitting a single \gls{MCE} is sufficient for producing consistent multiclass probabilistic estimates.

@@

		In this section we derive learning theoretic bounds that motivates our proposed hyperparameter learning algorithm, and discuss how it can be extended in various ways to enhance scalability and performance.

@@

		However, simply minimizing the empirical loss $\frac{1}{n} \sum_{i = 1}^{n} \mathcal{L}_{\epsilon}(y_{i}, \bvec{f}_{\theta, \lambda}(x_{i}))$ over the hyperparameters $(\theta, \lambda)$ could lead to an overfitted model. We therefore turn to using Rademacher complexity to control the model complexity of \glspl{MCE}.

		Let $\Theta$ and $\Lambda$ be a space of kernel hyperparameters and regularization parameters respectively. We define the class of \glspl{MCE} over these hyperparameter spaces by

		\begin{equation}
			F_{n}(\Theta, \Lambda) := \{ \bvec{f}_{\theta, \lambda}(x) : \theta \in \Theta, \lambda \in \Lambda \}.
			\label{eq:predictor_class_copy}
		\end{equation}

@@

		We denote $\hat{\mathcal{U}}^{(\theta, \lambda)}_{Y | X} \equiv W_{\theta, \lambda}^{T}$ to reflect the dependence on $(\theta, \lambda)$. We first restrict the space of hyperparameters by the norms of $W_{\theta, \lambda}$ and $k_{\theta}(x, x)$ to obtain an upper bound to the Rademacher complexity of $F_{n}(\Theta, \Lambda)$.

@@

		\cite{bartlett2002rademacher} showed that the expected risk can be bounded with high probability using the empirical risk and the Rademacher complexity of the loss composed with the function class. For a Lipchitz loss, \citet{ledoux2013probability} further showed that the latter can be bounded using the Rademacher complexity of the function class itself. We use these two results to arrive at the following probabilistic upper bound to our expected loss. We refer the reader to Appendix B for detailed discussion.

@@

			\begin{equation}
			\begin{aligned}
				&\mathbb{E}[\mathcal{L}_{e^{-1}}(Y, f(X))] \leq \frac{1}{n} \sum_{i = 1}^{n} \mathcal{L}_{\epsilon}(Y_{i}, f(X_{i})) \\
				&+ 4 e \; \alpha \rho + \sqrt{\frac{8}{n} \log{\frac{2}{\beta}}}.
			\end{aligned}
			\label{eq:specific_expected_loss_bound_for_multiclass_conditional_embedding_copy}
			\end{equation}

			See Lemma B.6 of Appendix B for proof.
		\end{theorem}

@@

		We can now apply \cref{thm:specific_expected_loss_bound_for_multiclass_conditional_embedding_copy} with the choice of $\rho = \| W_{\tilde{\theta}, \tilde{\lambda}} \|_{\mathrm{tr}}$ and $\alpha^{2} = \sup_{x \in \mathcal{X}} k_{\tilde{\theta}}(x, x)$. The probabilistic statement \eqref{eq:specific_expected_loss_bound_for_multiclass_conditional_embedding_copy} then only holds for $(\theta, \lambda) \in \Xi(\tilde{\theta}, \tilde{\lambda})$. However, we know that $(\tilde{\theta}, \tilde{\lambda}) \in \Xi(\tilde{\theta}, \tilde{\lambda})$, so we can choose $\theta \leftarrow \tilde{\theta}$ and $\lambda \leftarrow \tilde{\lambda}$. Since the only hyperparameters that appear now are $(\tilde{\theta}, \tilde{\lambda})$, we replace these symbols with a general $(\theta, \lambda)$ again to avoid cluttered notation. We now arrive at our final result.

@@

		In order for the bound \eqref{eq:expected_risk_bound_hyperparameter_learning_copy} to be non-trivial, we focus on bounded kernels in the sense that $\alpha^{2}(\theta) := \sup_{x \in \mathcal{X}} k_{\theta}(x, x) < \infty$. Gaussian kernels $k_{\theta}(x, x') = \sigma_{f}^{2} \exp{( - \frac{1}{2}(x - x') \Sigma^{-1} (x - x') )}$ and similar stationary kernels (e.g. Mat\'{e}rn) with sensitivity and length scales $\theta = (\sigma_{f}, \Sigma)$, for example, have $\alpha(\theta) = \sigma_{f}$. Since the training set itself is a sample of length $n$ drawn from $\mathbb{P}_{X Y}$, the inequality \eqref{eq:expected_risk_bound_hyperparameter_learning_copy} holds with probability $1 - \beta$ when the random variables $(X_{i}, Y_{i})$ are realized as the training observations $(x_{i}, y_{i})$. Motivated by this, we employ this upper bound as the learning objective for hyperparameter learning,
		\begin{equation}
			q(\theta, \lambda) := \frac{1}{n} \sum_{i = 1}^{n} \mathcal{L}_{\epsilon}(y_{i}, \bvec{f}_{\theta, \lambda}(x_{i})) + 4 e \; r(\theta, \lambda).
		\label{eq:learning_objective}
		\end{equation}

@@

		\paragraph{Batch Validation} While we simply instantiated $(X_{i}, Y_{i})$ to be the training observations in \cref{thm:expected_risk_bound_hyperparameter_learning_copy} to obtain \eqref{eq:learning_objective}, this does not have to be the case for batch updates. Instead, in each learning iteration, we could further split the batch into two sub-batches - one for training and one for validation. The training batch is used to form the \gls{MCE} $\bvec{f}_{\theta, \lambda}$ and \gls{RCB} $r(\theta, \lambda)$, while we evaluate the empirical risk on the validation batch,
		\begin{equation}
		q^{V}(\theta, \lambda) := \frac{1}{n_{V}} \sum_{i = 1}^{n_{V}} \mathcal{L}_{\epsilon}(y^{V}_{i}, \bvec{f}^{T}_{\theta, \lambda}(x^{V}_{i})) + \tau \; r^{T}(\theta, \lambda),
		\label{eq:validation_learning_objective}
		\end{equation}
		where $(T)$ and $(V)$ denotes training and validation. Note that in contrast to standard cross validation, not all data is required for each update due to the presence of the \gls{RCB}. Furthermore, although the multiplier on the \gls{RCB} is $4 e$, experiments show that generalization performance can improve if we uses a smaller multiplier $\tau < 4 e$. In practice, these two extensions work well together. Intuitively, by introducing a validation batch to measure empirical data fit, a smaller weight on the complexity penalty is required.

@@

		We developed a scalable hyperparameter learning framework for conditional mean embeddings with categorical targets based on Rademacher complexity bounds. These bounds reveal a novel data-dependent quantity $r(\theta, \lambda)$ which reflect the model complexity of \glspl{CME}, which we use as an regularization term in addition to the empirical loss for hyperparameter learning. Similar to a regularized least squares problem, it remains to be established what type of prior on the hyperparameters, if any, could correspond to such a regularizer, which would enable a Bayesian interpretation of our approach. We also envision that such a quantity could potentially be generalized to \glspl{CME} with arbitrary targets which would enable hyperparameter learning for general conditional mean embeddings in a way that is optimized for the loss relevant to the prediction task.

>>>>

\usepackage{glossaries}
\usepackage{acronym}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[colorlinks=true,linkcolor=black,citecolor=black]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

@@

		One notable success story in this domain are \gls{GPs}, which utilizes the marginal likelihood for hyperparameter learning. The marginal likelihood arises from its Bayesian formulation, and exhibits certain desirable properties -- in particular, the ability to automatically balance between data fit and model complexity. On the other hand, \glspl{CME} are not necessarily Bayesian, and hence they do not benefit from a natural marginal likelihood formulation, yet such a balance is critical when generalizing the model beyond known examples.

@@

		Let $p_{c}(x) := \mathbb{P}[Y = c | X = x]$ be the \textit{decision probability function} for class $c \in \mathbb{N}_{m}$, which is the probability of the class label $Y$ being $c$ when the example $X$ is $x$. Note that there are no restrictions on the input domain $\mathcal{X}$ as long as a kernel $k$ can be defined on it. For example, $x$ could be a continuous variable in $\mathcal{X} = \mathbb{R}^{d}$. We begin by writing this probability as an expectation of indicator functions,
		\begin{equation}
		p_{c}(x) := \mathbb{P}[Y = c | X = x] = \mathbb{E}[\mathbb{1}_{c}(Y) | X = x].
		\label{eq:decision_probability}
		\end{equation}	
		With $\mathbb{1}_{c} \in \mathcal{H}_{\delta}$, we let $g = \mathbb{1}_{c}$ in \eqref{eq:empirical_conditional_expectation} and $\bvec{1}_{c} := \{\mathbb{1}_{c}(y_{i})\}_{i = 1}^{n}$ to estimate the right hand side of \eqref{eq:decision_probability} by
		\begin{equation}
		\hat{p}_{c}(x) = f_{c}(x) := \bvec{1}_{c}^{T} (K + n \lambda I)^{-1} \bvec{k}(x).
		\label{eq:empirical_decision_probability}
		\end{equation}
		Let $\bvec{Y} := \begin{bmatrix} \bvec{1}_{1} & \bvec{1}_{2} & \cdots & \bvec{1}_{m} \end{bmatrix} \in \{0, 1\}^{n \times m}$ be the one hot encoded labels of $\{y_{i}\}_{i = 1}^{n}$. The vector of empirical decision probabilities over the classes $c \in \mathbb{N}_{m}$ is then
		\begin{equation}
		\hat{\bvec{p}}(x) = \bvec{f}(x) := \bvec{Y}^{T} (K + n \lambda I)^{-1} \bvec{k}(x) \in \mathbb{R}^{m}.
		\label{eq:empirical_decision_probability_vector}
		\end{equation}
		Since $\mathcal{U} = \hat{\mathcal{U}}_{Y | X}$ \eqref{eq:empirical_conditional_embedding} is the solution to a regularized least squares problem in the RKHS from $k(x, \cdot) \in \mathcal{H}_{k}$ to $l(y, \cdot) \in \mathcal{H}_{l}$ \citep{lever2012conditional}, \glspl{CME} can be seen as \glspl{KRR} in the \gls{RKHS}. In this case, because $\mathcal{Y} = \mathbb{N}_{m}$ is discrete, $\mathcal{H}_{\delta}$ can be identified with $\mathbb{R}^{m}$. As a result, for discrete $\mathcal{Y} = \mathbb{N}_{m}$ the rows of the \gls{MCE} can also be seen as $m$ \glspl{KRR} \citep{friedman2001elements} on binary $\{0, 1\}$-targets, where they all share the same input kernel. We now verify that the empirical decision probabilities \eqref{eq:empirical_decision_probability} do converge to the population decision probability.

@@

		This also allows \glspl{MCE} to be naturally applied to perform probabilistic classification in multiclass settings with categorical targets. In contrast, \glspl{SVC} do not naturally produce probabilities, while \glspl{GPC} require posterior approximations. Furthermore, multiclass extensions to \glspl{SVC} and \glspl{GPC} often employ the \gls{OVA} or \gls{OVO} scheme \citep{aly2005survey}, resulting in multiple separate binary classifiers. Instead, fitting a single \gls{MCE} is sufficient for producing consistent multiclass probabilistic estimates.

@@

		In this section we derive learning theoretic bounds that motivates our proposed hyperparameter learning algorithm, and discuss how it can be extended in various ways to enhance scalability and performance. From here onwards, we denote $\theta$ as the kernel hyperparameters of the kernel $k = k_{\theta}$.

@@

		However, simply minimizing the empirical loss $\frac{1}{n} \sum_{i = 1}^{n} \mathcal{L}_{\epsilon}(y_{i}, \bvec{f}_{\theta, \lambda}(x_{i}))$ over the hyperparameters $(\theta, \lambda)$ could lead to an overfitted model. We therefore employ Rademacher complexity bounds to control the model complexity of \glspl{MCE}.

		Let $\Theta$ and $\Lambda$ be a space of kernel hyperparameters and regularization parameters respectively. We define the class of \glspl{MCE} over these hyperparameter spaces by	
		\begin{equation}
			F_{n}(\Theta, \Lambda) := \{ \bvec{f}_{\theta, \lambda}(x) : \theta \in \Theta, \lambda \in \Lambda \}.
			\label{eq:predictor_class_copy}
		\end{equation}
		We denote $W_{\theta, \lambda}^{T} \equiv \hat{\mathcal{U}}^{(\theta, \lambda)}_{Y | X}$ and $\| W_{\theta, \lambda} \|_{\mathrm{tr}} = \| \hat{\mathcal{U}}^{(\theta, \lambda)}_{Y | X} \|_{HS}$ to reflect the dependence on $(\theta, \lambda)$ and for notational simplicity. We first restrict the space of hyperparameters by the norms of $W_{\theta, \lambda}$ and $k_{\theta}(x, x)$ to obtain an upper bound to the Rademacher complexity of $F_{n}(\Theta, \Lambda)$.

@@

@@

		\cite{bartlett2002rademacher} showed that the expected risk can be bounded with high probability using the empirical risk and the Rademacher complexity of the loss composed with the function class. For a Lipchitz loss, \citet{ledoux2013probability} further showed that the latter quantity can be bounded using the Rademacher complexity of the function class itself. We use these two results to arrive at the following probabilistic upper bound to our expected loss. We refer the reader to Appendix B for detailed discussion.

@@

			\begin{equation}
			\begin{aligned}
				&\mathbb{E}[\mathcal{L}_{e^{-1}}(Y, f(X))]  \\
				&\leq \frac{1}{n} \sum_{i = 1}^{n} \mathcal{L}_{\epsilon}(Y_{i}, f(X_{i})) + 4 e \; \alpha \rho + \sqrt{\frac{8}{n} \log{\frac{2}{\beta}}}.
			\end{aligned}
			\label{eq:specific_expected_loss_bound_for_multiclass_conditional_embedding_copy}
			\end{equation}

			See Theorem B.6 of Appendix B for proof.
		\end{theorem}

@@

		We can now choose $\tilde{\theta} \in \Theta$, $\tilde{\lambda} \in \Lambda$ and apply \cref{thm:specific_expected_loss_bound_for_multiclass_conditional_embedding_copy} with the choice of $\rho = \| W_{\tilde{\theta}, \tilde{\lambda}} \|_{\mathrm{tr}}$ and $\alpha^{2} = \sup_{x \in \mathcal{X}} k_{\tilde{\theta}}(x, x)$ and by considering only the hyperparameters $(\theta, \lambda) \in \Xi(\tilde{\theta}, \tilde{\lambda})$. The probabilistic statement \eqref{eq:specific_expected_loss_bound_for_multiclass_conditional_embedding_copy} then only holds for $(\theta, \lambda)$ within $\Xi(\tilde{\theta}, \tilde{\lambda})$. In particular, since $(\tilde{\theta}, \tilde{\lambda}) \in \Xi(\tilde{\theta}, \tilde{\lambda})$, it holds for $\theta = \tilde{\theta}$ and $\lambda = \tilde{\lambda}$. Applying this choice, the only variables that remain now are $(\tilde{\theta}, \tilde{\lambda})$. We then replace these symbols with a general $(\theta, \lambda)$ again to avoid cluttered notation. We now arrive at our final result.

@@

		% In order for the bound \eqref{eq:expected_risk_bound_hyperparameter_learning_copy} to be non-trivial, we focus on bounded kernels in the sense that $\alpha^{2}(\theta) := \sup_{x \in \mathcal{X}} k_{\theta}(x, x) < \infty$. Gaussian kernels $k_{\theta}(x, x') = \sigma_{f}^{2} \exp{( - \frac{1}{2}(x - x') \Sigma^{-1} (x - x') )}$ and similar stationary kernels (e.g. Mat\'{e}rn) with sensitivity and length scales $\theta = (\sigma_{f}, \Sigma)$, for example, have $\alpha(\theta) = \sigma_{f}$.
		In particular, $r(\theta, \lambda)$ is an upper bound to the Rademacher complexity of a relevant class of \glspl{MCE} based on the hyperparameters $\Xi(\theta, \lambda)$. We call $r(\theta, \lambda)$ the \gls{RCB} and use it to measure the model complexity of a \gls{MCE} with hyperparameters $(\theta, \lambda)$. Since the training set itself is a sample of length $n$ drawn from $\mathbb{P}_{X Y}$, the inequality \eqref{eq:expected_risk_bound_hyperparameter_learning_copy} holds with probability $1 - \beta$ when the random variables $(X_{i}, Y_{i})$ are realized as the training observations $(x_{i}, y_{i})$. Motivated by this, we employ this upper bound as the learning objective for hyperparameter learning,
		\begin{equation}
			q(\theta, \lambda) := \frac{1}{n} \sum_{i = 1}^{n} \mathcal{L}_{\epsilon}(y_{i}, \bvec{f}_{\theta, \lambda}(x_{i})) + 4 e \; r(\theta, \lambda).
		\label{eq:learning_objective}
		\end{equation}

		In particular, the first term is an empirical risk that measures data fit, and the second term is the \gls{RCB} that measures model complexity.

@@

		\paragraph{Batch Validation} While we simply instantiated $(X_{i}, Y_{i})$ to be the training observations in \cref{thm:expected_risk_bound_hyperparameter_learning_copy} to obtain \eqref{eq:learning_objective}, this does not have to be the case for batch updates. Instead, in each learning iteration, we could further split the batch into two sub-batches -- one for training and one for validation. The training batch is used to form the \gls{MCE} $\bvec{f}_{\theta, \lambda}$ and \gls{RCB} $r(\theta, \lambda)$, while we evaluate the empirical risk on the validation batch,
		\begin{equation}
		q^{V}(\theta, \lambda) := \frac{1}{n_{V}} \sum_{i = 1}^{n_{V}} \mathcal{L}_{\epsilon}(y^{V}_{i}, \bvec{f}^{T}_{\theta, \lambda}(x^{V}_{i})) + \tau \; r^{T}(\theta, \lambda),
		\label{eq:validation_learning_objective}
		\end{equation}
		where $(T)$ and $(V)$ denotes training and validation. Note that in contrast to standard cross validation, not all data is required for each update due to the presence of the \gls{RCB}. Furthermore, although the multiplier on the \gls{RCB} is $4 e$, experiments show that generalization performance can improve if we uses a smaller multiplier $\tau < 4 e$. In practice, these two extensions work well together. Intuitively, by introducing a validation batch to measure empirical data fit, a smaller weight on the complexity penalty is required.

@@

		We developed a scalable hyperparameter learning framework for \glspl{CME} with categorical targets based on Rademacher complexity bounds. These bounds reveal a novel data-dependent quantity $r(\theta, \lambda)$ which reflect its model complexity, which we use as an regularization term in addition to the empirical loss for hyperparameter learning. Similar to a regularized least squares problem, it remains to be established what type of prior on the hyperparameters, if any, could correspond to such a regularizer, resulting in a Bayesian interpretation of our approach. We also envision that such a quantity could potentially be generalized to \glspl{CME} with arbitrary targets, which would enable hyperparameter learning for general conditional mean embeddings in a way that is optimized for the prediction task.
